The watershed transform in ITK - discussion and new developments

### 2006-04-23 article/Article.tex
add implementation section
darcs-hash:20060423153043-2fc9d-8b6205fc26a93aa83db676431c1013a53b69e72e.gz

\subsubsection{Algorithm 2}

@@

\subsubsection{Algorithm 3}

@@

Contrary to the previous algorithm, the label in the output image
can't be used to determine if the pixel has already been added to
the queue, and so we need to maintain this status by another way.
It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
in ??? how this status can be implemented.

@@

@@

\section{Timing and performance}
streaming and threading.

>>>>

\subsubsection{Algorithm 2}
\label{sect:BeucherAlgorithm}

@@

\subsubsection{Algorithm 3}
\label{sect:MeyerAlgorithm}

@@

Contrary to the previous algorithm, the label in the output image
can't be used to determine if the pixel has already been added to
the queue, and so we need to maintain this status by another way.
It is the role of $S$ in Figure \ref{MeyerAlgorithm}. We'll discuss
in \ref{sect:implementation} how this status can be implemented.

@@

\section{Implementation}
\label{sect:implementation}

We provide with this article two new classes:

  \subsection{MorphologicalWatershedFromMarkersImageFilter}

\verb$MorphologicalWatershedFromMarkersImageFilter$ implement the
algorithms \ref{sect:BeucherAlgorithm} and \ref{sect:MeyerAlgorithm}.
They are accessible respectively when the attribute \verb$MarkWatershed$
is \verb$false$ or is \verb$true$. The filter is templated over the
input image type, which can be any scalar type, and the label image type
which should be an integer. The filter take 2 inputs:
\begin{itemize}
  \item a grayscale image which can be set with \verb$SetInput(image)$,
\verb$SetInput(0, image)$ or \verb$SetInput1(image)$.
  \item a marker image which can be set with \verb$SetMarkerImage(image)$
and \verb$SetInput2(image)$. \verb$SetInput(0, image)$ should be avoided
because it will fail if the input image and the marker image are not of the
same type. The markers are all the non-zero values in the image.
\end{itemize}

The algorithm \ref{sect:MeyerAlgorithm} require to keep the status of all
the pixels: $already added to the queue$ or $not yet added to the queue$.
This task is often done by using negative values in the output image
for the pixels not yet added to the queue. In our implementation, we have
chosen to use a binary image to store this, while labels always take values
above 0 and watershed lines are marked with value 0. The negative values
option have several drawbacks:
\begin{itemize}
  \item it requires to use a different image type than the label type defined
by the user in case the user have chosen an unsigned pixel type.
  \item it requires to use one of the value which can be possibly aready used
by a label in the marker image, in case of a signed type.
\end{itemize}
Those problems can be avoided\footnote{It's not true for the unsigned long
pixel type} by using a type which contains a larger range of value than the
type choosed by the user, but then, we must copy the data to an image of
the type chosen by the user at the end of the algorithm.

We prefered using a separated binary image (of type \verb$itk::Image<bool, dim>$)
to store this information. This solution avoid the need to copy the image at
the end, and so should be more efficient in computation time, and require only
one bit per pixel and so should be more memory efficient as well.

  \subsection{MorphologicalWatershedImageFilter}

\verb$MorphologicalWatershedImageFilter$ perform a watershed transformation
without marker. It encapsulate several filters:
\begin{itemize}
  \item \verb$HMinimaImageFilter$
  \item \verb$RegionalMinimaImageFilter$
  \item \verb$ConnectedComponentImageFilter$
  \item \verb$MorphologicalWatershedFromMarkersImageFilter$
\end{itemize}
While \verb$RegionalMinimaImageFilter$, \verb$ConnectedComponentImageFilter$,
and \verb$MorphologicalWatershedFromMarkersImageFilter$ are the usal steps
of the waterhed usage, the \verb$HMinimaImageFilter$ is a convenient way
to avoid the classical watershed's over segmentation by allowing the user
to select the dynamic of the minima.

\section{Timing and performance}

streaming and threading.

@@
